---
title: |
  | Introduction to Data Science
  | R Tutorial
  | Week 7 
author: |
  | Ilse Cuevas Andrade and Dr. Gah-Yi Ban
  | Imperial College Business School

output:
  pdf_document: default
  html_document:
    df_print: paged
---



## **Exercise: Detecting Outliers in `airquality` Dataset**

This exercise guides you through detecting outliers in the `Ozone` column of the `airquality` dataset using a boxplot, IQR method, and a scatterplot for visualisation.

---

### **Step 1: Load and Explore the Dataset**

We will use the `airquality` dataset, which contains air quality measurements. Start by loading and inspecting the data.

```{r}
# Load tidyverse
library(tidyverse)
```


```{r}
# Load the airquality dataset and inspect it
data("airquality")


head(airquality)


airquality<-airquality %>%
  rowid_to_column("Index") %>%  # Add an index column for plotting
  drop_na(Ozone)               # Remove rows with missing Ozone values


```



*Exercise 1*
Create a boxplot of the Ozone column to visually inspect for outliers.

```{r}
# Boxplot of Ozone levels

```


*Exercise 2*
Follow these steps to calculate and use the IQR to detect outliers:

Compute the quartiles (Q1, Q3) and the IQR.
Calculate the lower and upper bounds for detecting outliers.
Identify which data points fall outside these bounds.

```{r}
# Calculate IQR and classify points as outliers
ozone_with_outliers <- ???

# Display the bounds and outliers
ozone_bounds <- ???

cat("Lower and Upper Bounds for Ozone Levels:\n")
print(ozone_bounds)

cat("\nOutliers in Ozone Levels:\n")
print(ozone_with_outliers %>% filter(Is_Outlier == "Outlier") %>% select(Index, Ozone))


```



*Exercise 3*
Create a scatterplot where outliers are highlighted in red.
```{r}
# Scatterplot of Ozone levels with outliers highlighted


```





## The end
