---
title: |
  | Introduction to Data Science
  | R Tutorial part 2
  | Week 5  
author: |
  | Ilse Cuevas Andrade and Dr. Gah-Yi Ban
  | Imperial College Business School
  
output:
  pdf_document: default
  html_document:
    df_print: paged
---

\newpage
\tableofcontents
\listoffigures
\listoftables
\newpage


Getting Started with R Notebook
======================

This is an *R Notebook*.

**NOTE** The lines before line 23 are the settings of this R Notebook document.
You can alter the text in blue colour. Please do not change anything else within
the first 22 lines unless you are going to deal with some extra challenges.

knit to PDF
----------

Use the button "Knit" and select the "Knit to PDF" option from the drop-down menu
to knit this document to a PDF file. You can keep doing this step when you are
updating your R Notebook to view the results in the output PDF file.

An example code
-----------

Complete the follow task by adding a code chunk which contains the following elements:

- code chunk name

- variables

- arithmetic calculation

- comments

Improve presentation of PDF
-----------

The following paragraphs summarize R Markdown key features.

## R Markdown

The special characters used to format this document come from a language called
*Markdown*. Now that you've created a professional-looking copy of the handout,
let's learn about the different characters we used for formatting it.

### Headings

Up until now, we created headings by using a dashed line like so `----------`.
A more common way to specify headings in Markdown is with the hash # symbol. Do
not confuse this with comments in R - these only occur when inside a code chunk!

##### Heading Levels

There are six levels of headings specified by the hash symbol. Top-level
headings are specified by # (one hash symbol) , where the lowest level headings
are specified by ###### (six hash symbols).

### Emphasis

You can make text *italic* by enclosing words or phrases in *single asterisks* *

Use **double asterisks** to make text **bold** **

Code is often represented in `monospace text`, enclosed in `single backticks` `

You can also ~~strike text out~~ with double tildes ~~

### Lists

Dashes are used to create lists. For nesting items within a list, indent the
nested items by a consistent amount (2-spaces are recommended). Ensure there is
a blank line between paragraphs and lists in the document.

- first item
  - nested item
    - further nesting
- second item

### Code

You have seen how we represent larger blocks of code in code chunks, starting
and ending with three single backticks.

```
Generic code chunks can contain code in any language, but will not be evaluated.
```

If we want code blocks to be evaluated with R, we use the {r} directive,
optionally giving the code chunk a name.
```{r Code chunk example}
print("This code will appear in the rendering and it will be evaluated")
```

We can specify several options inside the {r} directive, separated by commas. To
force the code chunk to not evaluate and not show a result, use `eval=FALSE`
```{r Non-evaluated code chunk, eval=FALSE}
print("This code will appear in the rendering but will NOT be evaluated")
```

Other times, we want to see the result, but not the code that produced it. For
this, use `echo=FALSE`
```{r Hidden code chunk, echo=FALSE}
print("Only the result of this code chunk will appear in the rendered document")
```


[Markdown](https://rmarkdown.rstudio.com/lesson-1.html) allows you to compose
and improve the presentation of final output in RStudio. Click on the link 
https://rmarkdown.rstudio.com/lesson-1.html
for 
further information and try to improve the presentation of this document in PDF
format. For example,

- Add more chapters/sections.

- Add bullet points.

- Add highlights, such as, bold, italic fonts etc.

- Remove unnecessary list from the table of contents.

**Hint**: edit the lines within the first 20 lines as needed, then knit again.

## The End

